---
title: "Descriptive Statistics of Demographic Information  \nfor Healthy Participants"
author: "Marielena"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::pdf_document:  
    fig_caption: yes        
---

```{r setup, include = FALSE}
## R-setup

library(tidyverse)
library(readxl)
library(ggpubr)

## import data
getwd() 
setwd("../Demo") ## set folder 'Demo' as the working directory
healthy_all <- readxl::read_xlsx("2023.07.24.Demo-Healthy.xlsx")

## remove participants with MoCA score <25

healthy <- healthy_all[which(healthy_all$moca> 25), ] ## should remove 19 participants 
```

## Number of Participants

```{r number of p}
nrow(healthy)#number of participants
```

```{r number of f}
nrow(healthy[healthy$sex == "f",]) #number of women
```

```{r number of m}
sum(with(healthy,sex == "m")) #number of men
```

## Descriptive Statistics (Age)

```{r descriptive age}
mean(healthy$age) #mean
```

```{r range}
range(healthy$age) #min & max
```

```{r sd}
sd(healthy$age) #standard-deviation
```

```{r var}
var(healthy$age) #variance
```

 

 

 

## Descriptive Statistics (Education)

```{r descriptive edu mean}
mean(healthy$eduyears)#mean
```

```{r descriptive edu range}
range(healthy$eduyears) #min & max
```

```{r descriptive edu sd}
sd(healthy$eduyears) #standard-deviation 
```

```{r descriptive edu var}
var(healthy$eduyears) #variance
```

## I. Age Histograms

```{r hist, echo=FALSE}

breakshist1=seq(18,85,by=1)
#length(breakshist1)

agehist_1 <- hist(healthy$age, 
     main="Histogram of healhty participants' age",
     sub="per year",
     xlab="age",
     xaxt="n",
     ylab="number of participants",
     ylim=c(0,25),
     border="blue", 
     col="green",
     las=1, 
     breaks=breakshist1,
     xaxp=c(18,85, 1))

#print(agehist_1)
axis(side=1, at=seq(18,85, by=1), labels=seq(18,85, 1), cex.axis=0.50)

```

### Number of participants per age

```{r, echo=FALSE}
agg_tbl1 <- healthy %>% group_by(age) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
df1 <- agg_tbl1 %>% as.data.frame()
df1

```

### a. 10-year age cohorts

```{r hist10, echo=FALSE}

agehist_10 <- hist(healthy$age, 
     main="Histogram of healhty participants' age", 
     sub="per decade",
     xlab="age", 
     ylab="number of participants",
     border="blue", 
     col="green",
     xlim=c(0,100),
     labels = NULL,
     ylim=c(0,60),
     las=1, 
     breaks=seq(10,90, 10))

#print(agehist_10)
axis(side=1,at=seq(10,100,10),labels=seq(10,100,10))

```

```{r hist10 paper, echo=FALSE}

agehist_10b <- hist(healthy$age,
     main=NULL,
     xlab="age", 
     ylab="frequency",
     border="black", 
     col="gray",
     xlim=c(10,90),
     labels = NULL,
     ylim=c(0,60),
     las=1, 
     breaks=seq(10,90, 10))

#print(agehist_10)
axis(side=1,at=seq(10,100,10),labels=seq(10,100,10))

```

### Number of Participants per life decade

```{r group participant per decade, echo=FALSE}
## Group by count using dplyr
agg_tbl <- healthy %>% group_by(age = age - age %% 10) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
#agg_tbl

## Convert tibble to df
df10 <- agg_tbl %>% as.data.frame()
df10
```

### b. 5-year age cohorts

```{r hist5, echo=FALSE}
g = healthy$age
m<-mean(g)
std<-sd(g)

myhist <- hist(g,
     main="Histogram of healhty participants' age", 
     sub="per five-year intervals",
     xlab="age", 
     ylab="number of participants",
     border="darkgray", 
     col="darkmagenta",
     xaxt='n',
     ylim=c(0,40),
     las=1, 
     breaks=14)

multiplier <- myhist$counts / myhist$density
mydensity <- density(g)
mydensity$y <- mydensity$y * multiplier[1]

myhist <- hist(g,
     main="Histogram of healhty participants' age", 
     sub="per five-year intervals",
     xlab="age", 
     ylab="number of participants",
     border="darkgray", 
     col="darkmagenta",
     xaxt='n',
     ylim=c(0,40),
     las=1, 
     breaks=14)
  lines(mydensity)
     myx <- seq(min(g), max(g), length.out= 100)
     mymean <- mean(g)
     mysd <- sd(g)
normal <- dnorm(x = myx, mean = mymean, sd = mysd)
     lines(myx, normal * multiplier[1], col = "blue", lwd = 2)
sd_x <- seq(mymean - 3 * mysd, mymean + 3 * mysd, by = mysd)
sd_y <- dnorm(x = sd_x, mean = mymean, sd = mysd) * multiplier[1]
segments(x0 = sd_x, y0= 0, x1 = sd_x, y1 = sd_y, col = "firebrick4", lwd = 2)
axis(side=1, at=seq(10,85, 5), labels=seq(10,85, 5))
plot=TRUE

```

### Number of Participants per five-year age groups

```{r group participant per five-year intervals, echo=FALSE}
## Group by count using dplyr
agg_tbl2 <- healthy %>% group_by(age = age - age %% 5) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
#agg_tbl

## Convert tibble to df
df5 <- agg_tbl2 %>% as.data.frame()
df5
```

---

## II. Education histogram

```{r hist-edu, echo=FALSE}

eduhist <- hist(healthy$eduyears, 
     main="Histogram for healhty participants' education", 
     xlab="years of official education", 
     ylab="participants",
     border="red", 
     col="blue",
     xaxt='n',
     xlim=c(5,28),
     ylim=c(0,85),
     las=1, 
     breaks="FD")
axis(side=1, at=seq(6,26, 2), labels=seq(6,26, 2))

print(eduhist)
```

# Age - Edu Scatter plot

```{r scatterplot age$edu, echo = FALSE}
x <- healthy$age
y <- healthy$eduyears
## Plot with main and axis titles
## Change point shape (pch = 19) and remove frame.
plot(x, y, main = "Participants' Age - Years of education",
     xlab = "age", ylab = "years of education",
     pch = 19, frame = FALSE)

## Add regression line
plot(x, y, main = "Participants' Age - Years of education",
     xlab = "age", ylab = "years of education",
     pch = 19, frame = FALSE)
abline(lm(y ~ x, data = mtcars), col = "blue")
```

## Correlations

### a. Spearman Correlation

```{r spearman corr age-edu, echo=FALSE}
cor(healthy$eduyears, healthy$age, method = c ("spearman"))


ggscatter(healthy, x = "eduyears", y = "age", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "spearman",
          xlab = "years of official education", ylab = "age")
```


### b. Pearson Correlation

```{r pearson corr age-edu, echo=FALSE}

cor(healthy$eduyears, healthy$age, method = c ("pearson"))

library("ggpubr")
ggscatter(healthy, x = "eduyears", y = "age", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "years of official education", ylab = "age")
```
